---
title: "Analysis"
---



## Key Takeaways

-   Internal migration patterns in Türkiye reveal significant regional differences in population movement.
-   Western regions such as Marmara, Ege, and Akdeniz generally experience population gain, while eastern regions tend to experience population loss.
-   Migration decisions are strongly associated with family-related reasons, better living conditions, education, and employment opportunities.
-   Regional migration analysis provides valuable insight into socio-economic inequalities and demographic changes across the country.

-----------------------------------------------------------------------


## Regional Balance of Internal Migration

Migration patterns in Turkey are not evenly distributed across geographical regions. Some regions attract population through economic opportunities, urban development, education, and industrial activity, while others experience population loss due to outward migration.

To better understand these regional dynamics, provinces were grouped according to Turkey’s seven geographical regions. Total incoming migration and outgoing migration values were calculated for each region, and net migration values were obtained by subtracting outgoing migration from incoming migration.

Positive values indicate that a region receives more people than it loses, whereas negative values indicate population loss through migration.

The graph below compares the regional balance of internal migration and highlights which regions function as migration attractors and which regions experience migration decline.
```{r}
#| echo: FALSE
#| warning: FALSE
#| message: FALSE

library(tidyverse)
library(readr)

final_data_clean <- read_csv(
  "final_data_clean.csv",
  locale = locale(encoding = "UTF-8"),
  show_col_types = FALSE
)
```

```{r}
#| echo: FALSE
#| warning: FALSE
#| message: FALSE
#| fig-width: 11
#| fig-height: 7
#| out-width: 100%

library(dplyr)
library(ggplot2)
library(stringi)
library(scales)

options(scipen = 999)

marmara <- c("istanbul","edirne","kirklareli","tekirdag","canakkale",
             "kocaeli","yalova","sakarya","bilecik","bursa","balikesir")

ege <- c("izmir","manisa","aydin","denizli","mugla",
         "afyonkarahisar","kutahya","usak")

akdeniz <- c("antalya","isparta","burdur","adana",
              "mersin","hatay","osmaniye","kahramanmaras")

ic_anadolu <- c("ankara","konya","karaman","aksaray","nigde",
                 "nevsehir","kirsehir","kirikkale","yozgat",
                 "kayseri","sivas","cankiri","eskisehir")

karadeniz <- c("artvin","rize","bayburt","trabzon","gumushane",
                "giresun","ordu","tokat","amasya","samsun",
                "sinop","corum","kastamonu","bartin",
                "karabuk","zonguldak","duzce","bolu")

dogu_anadolu <- c("erzurum","erzincan","kars","tunceli",
                   "bingol","elazig","malatya","hakkari",
                   "van","bitlis","mus","agri","igdir","ardahan")

guneydogu <- c("gaziantep","kilis","adiyaman","sanliurfa",
                "diyarbakir","mardin","batman","siirt","sirnak")

regional_data <- final_data_clean %>%
  
  mutate(
    city = stri_trans_general(il, "Latin-ASCII"),
    city = tolower(city)
  ) %>%
  
  mutate(
    region = case_when(
      city %in% marmara ~ "Marmara",
      city %in% ege ~ "Aegean",
      city %in% akdeniz ~ "Mediterranean",
      city %in% ic_anadolu ~ "Central Anatolia",
      city %in% karadeniz ~ "Black Sea",
      city %in% dogu_anadolu ~ "Eastern Anatolia",
      city %in% guneydogu ~ "Southeastern Anatolia",
      TRUE ~ "Other"
    )
  ) %>%
  
  filter(region != "Other") %>%
  
  group_by(region) %>%
  
  summarise(
    incoming = sum(gelen_sayi, na.rm = TRUE),
    outgoing = sum(giden_sayi, na.rm = TRUE),
    net_migration = incoming - outgoing,
    .groups = "drop"
  ) %>%
  
  arrange(desc(net_migration))

ggplot(
  regional_data,
  aes(
    x = reorder(region, net_migration),
    y = net_migration,
    fill = net_migration > 0
  )
) +
  
  geom_col(width = 0.7) +
  
  coord_flip() +
  
  geom_text(
    aes(
      label = paste0(
        ifelse(net_migration > 0, "+", ""),
        round(net_migration / 1000, 1),
        " K"
      )
    ),
    
    hjust = ifelse(regional_data$net_migration > 0, -0.2, 1.2),
    
    size = 4,
    
    fontface = "bold",
    
    color = "#444444"
  ) +
  
  scale_fill_manual(
    values = c(
      "TRUE" = "#003366",
      "FALSE" = "#f1c40f"
    ),
    
    labels = c(
      "FALSE" = "Net Loss",
      "TRUE" = "Net Gain"
    ),
    
    name = NULL
  ) +
  
  scale_y_continuous(
    labels = label_number(
      suffix = " K",
      scale = 1e-3
    ),
    
    expand = expansion(mult = c(0.15, 0.15))
  ) +
  
  labs(
    title = "Regional Balance of Internal Migration",
    
    subtitle = "Population gains and losses across Turkey's geographical regions",
    
    x = NULL,
    
    y = "Net Migration"
  ) +
  
  theme_minimal(base_family = "sans") +
  
  theme(
    
    plot.title = element_text(
      face = "bold",
      size = 20,
      color = "#003366"
    ),
    
    plot.subtitle = element_text(
      size = 12,
      color = "#666666",
      margin = margin(b = 15)
    ),
    
    axis.text.y = element_text(
      size = 11,
      face = "bold",
      color = "#444444"
    ),
    
    axis.text.x = element_text(
      color = "#555555"
    ),
    
    axis.title = element_text(
      face = "bold",
      color = "#003366"
    ),
    
    panel.grid.major.y = element_blank(),
    
    panel.grid.minor = element_blank(),
    
    legend.position = "top",
    
    legend.text = element_text(
      size = 10,
      face = "bold"
    ),
    
    plot.margin = margin(
      30,
      40,
      30,
      40
    )
  )
```

The visualization clearly demonstrates that migration balance differs significantly between regions. Regions such as Marmara and the Aegean show positive net migration values, meaning that they attract more people than they lose. These regions are generally associated with stronger economies, industrial development, educational institutions, and urban opportunities.

In contrast, regions such as Eastern Anatolia and Southeastern Anatolia experience negative net migration values, indicating higher levels of outward migration. This pattern may reflect differences in employment opportunities, living standards, infrastructure, and regional development.

Overall, the analysis highlights the unequal geographical distribution of migration in Turkey and shows how migration patterns are closely connected to socio-economic conditions and regional attractiveness.



------------------------------------------------------------------------

## Total Internal Migration Flow by Province

While regional analysis provides a broader understanding of migration dynamics, examining migration flow at the provincial level allows us to identify which specific cities dominate internal migration in Türkiye.

In this visualization, provinces are ranked according to their total migration flow volume. The total flow value represents the combined amount of incoming and outgoing migration activity for each province. This approach highlights provinces with the highest population mobility regardless of whether they experience net migration gain or loss.

The graph reveals a highly uneven distribution of migration flow across provinces. A limited number of metropolitan cities stand out with significantly higher migration volumes compared to the rest of the country. Provinces such as İstanbul, Ankara, and İzmir clearly dominate the migration network due to their economic activity, educational opportunities, transportation accessibility, and urban infrastructure.

Another important observation is the strong concentration of migration in major urban centers. Most provinces remain clustered at relatively lower migration levels, while a small number of highly developed provinces create a sharp increase at the upper end of the distribution. This indicates that internal migration in Türkiye is heavily centralized around a few key metropolitan areas.

The visualization also demonstrates the presence of outlier provinces with exceptionally high migration activity. These outliers suggest that migration patterns are not evenly spread across the country and are instead shaped by regional economic attraction, labor markets, and quality-of-life differences.

```{r}
#| echo: TRUE
install.packages(c("dplyr", "ggplot2", "scales"))

library(dplyr)
library(ggplot2)
library(scales)

options(scipen = 999)

il_akis_sirali <- final_data_clean %>%
  group_by(il) %>%
  summarise(
    total_flow = sum(gelen_sayi, na.rm = TRUE) + sum(giden_sayi, na.rm = TRUE)
  ) %>%
  ungroup() %>%
  arrange(total_flow)

ggplot(il_akis_sirali, aes(x = reorder(il, total_flow), y = total_flow)) +
  geom_segment(aes(x = reorder(il, total_flow), xend = il, y = 0, yend = total_flow), 
               color = "#d6d6d6", size = 0.4) +
  geom_point(color = "#003366", 
             size = 2.5, 
             shape = 16, 
             alpha = 1) +
  scale_y_continuous(
    labels = label_number(suffix = " K", scale = 1e-3),
    breaks = seq(0, max(il_akis_sirali$total_flow) + 100000, by = 100000),
    expand = expansion(mult = c(0, 0.12))
  ) +
  theme_minimal(base_family = "sans") +
  labs(
    title = "Total Internal Migration Flow by Province",
    subtitle = "Provinces ranked by volume of the flow",
    x = NULL, 
    y = "Total Flow"
  ) +
  theme(
    plot.background = element_rect(fill = "white", color = NA),
    panel.background = element_rect(fill = "white", color = NA),
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, size = 7.5, 
                               color = "#5a5a5a", face = "bold"),
    axis.text.y = element_text(size = 9, color = "#5a5a5a", face = "bold"),
    panel.grid.major.y = element_line(color = "#e0e0e0", size = 0.4), 
    panel.grid.major.x = element_blank(),
    panel.grid.minor = element_blank(),
    plot.title = element_text(face = "bold", size = 20, color = "#003366", hjust = 0),
    plot.subtitle = element_text(size = 12, color = "#7a7a7a", margin = margin(b = 20), hjust = 0),
    axis.line.x = element_line(color = "#cccccc", size = 0.6),
    plot.margin = margin(30, 30, 30, 30)
  )
```

------------------------------------------------------------------------

## In-Migration Map of Türkiye

This map visualizes in-migration patterns across Türkiye at the provincial level. Each point represents a province, and the size of the point shows the total number of people migrating into that province. Larger points indicate provinces that receive higher levels of in-migration.

The color of each point represents the most common reason for in-migration in that province. This allows us to observe not only where people move, but also why they move. Reasons such as education, better housing and living conditions, family-related migration, and returning to hometowns help explain the social and economic motivations behind internal migration.

This visualization is useful because it combines geographical location, migration volume, and migration reason in a single map. It shows that migration is not randomly distributed across Türkiye. Instead, some provinces attract more people due to opportunities related to education, employment, family networks, and urban living standards.

```{r}
#| echo: FALSE
#| warning: FALSE
#| message: FALSE
#| fig-width: 12
#| fig-height: 8

library(dplyr)
library(ggplot2)
library(sf)
library(rnaturalearth)
library(stringr)
library(stringi)
library(scales)

options(scipen = 999)

il_ozet <- final_data_clean %>%
  mutate(
    clean_neden = tolower(stri_trans_general(neden, "Latin-ASCII")),
    
    standart_neden = case_when(
      str_detect(clean_neden, "egitim|education") ~ "Education",
      
      str_detect(clean_neden, "konut|yasam|housing|living") ~ "Better Living",
      
      str_detect(clean_neden, "memleket|geri donme|returning") ~ "Returning Home",
      
      str_detect(clean_neden, "bagimli|hane|household|family") ~ "Family Related",
      
      TRUE ~ "Other"
    )
  ) %>%
  
  group_by(il) %>%
  
  mutate(total_in_migration = sum(gelen_sayi, na.rm = TRUE)) %>%
  
  filter(gelen_sayi == max(gelen_sayi, na.rm = TRUE)) %>%
  
  slice(1) %>%
  
  select(
    il,
    top_reason = standart_neden,
    total_in_migration
  ) %>%
  
  ungroup()

clean_city_name <- function(x) {
  x %>%
    as.character() %>%
    stri_trans_general("Latin-ASCII") %>%
    tolower() %>%
    str_trim() %>%
    str_replace_all("ı", "i") %>%
    str_replace_all("^afyon$", "afyonkarahisar") %>%
    str_replace_all("^maras$", "kahramanmaras")
}

turkey_map <- ne_states(
  country = "Turkey",
  returnclass = "sf"
)

map_data <- turkey_map %>%
  
  mutate(match_name = clean_city_name(name)) %>%
  
  left_join(
    il_ozet %>%
      mutate(match_name = clean_city_name(il)),
    by = "match_name"
  )

map_points <- st_point_on_surface(map_data)

ggplot() +
  
  geom_sf(
    data = turkey_map,
    fill = "gray95",
    color = "gray70",
    linewidth = 0.25
  ) +
  
  geom_sf(
    data = map_points,
    aes(
      size = total_in_migration,
      color = top_reason
    ),
    alpha = 0.85
  ) +
  
  scale_size_area(
    max_size = 14,
    name = "Total In-Migration",
    labels = comma
  ) +
  
  scale_color_manual(
    values = c(
      "Education" = "#003366",
      "Better Living" = "#2ca25f",
      "Returning Home" = "#f1c40f",
      "Family Related" = "#5dade2",
      "Other" = "gray60"
    ),
    
    na.value = "gray70",
    
    name = "Top Reason for In-Migration"
  ) +
  
  labs(
  title = "Internal Migration Map of Turkey",
  subtitle = "Top reasons for in-migration and total in-migration    volumes by province"
  ) +
  
  theme_void() +
  
  theme(
    plot.title = element_text(
      face = "bold",
      size = 20,
      color = "#003366",
      hjust = 0.5
    ),
    
    plot.subtitle = element_text(
      size = 12,
      color = "#555555",
      hjust = 0.5,
      margin = margin(b = 15)
    ),
    
    legend.position = "bottom",
    
    legend.box = "vertical",
    
    legend.title = element_text(
      face = "bold",
      size = 10
    ),
    
    legend.text = element_text(size = 9),
    
    plot.margin = margin(10, 10, 10, 10)
  ) +
  
  guides(
    color = guide_legend(
      ncol = 3,
      title.position = "top"
    ),
    
    size = guide_legend(
      title.position = "top"
    )
  )
```

The map shows that provinces with larger in-migration volumes are visually concentrated through larger points. These provinces can be interpreted as stronger migration attractors within Türkiye. The colors also reveal that the dominant reasons for migration differ across provinces. For example, some provinces mainly attract people because of education, while others are more related to family movement or better living conditions.

Overall, this map supports the idea that internal migration is shaped by both geographical and socio-economic factors. It also helps present migration patterns in a more intuitive way by connecting statistical values with their spatial distribution.

------------------------------------------------------------------------

## Out-Migration Map of Turkey

While the previous visualization focused on incoming migration patterns, this map examines the distribution of outgoing migration across provinces in Turkey. The purpose of this analysis is to identify which provinces experience the highest levels of population outflow and to explore the dominant reasons behind these migration movements.

Each point on the map represents a province. The size of the point corresponds to the total amount of outgoing migration from that province. Larger circles indicate provinces with higher out-migration volumes.

In addition, colors are used to represent the most common reason for out-migration in each province. This provides a more comprehensive understanding of migration behavior by combining migration intensity with migration motivation.

The visualization highlights how migration dynamics differ geographically and demonstrates that population loss is not evenly distributed across the country.

```{r}
#| echo: FALSE
#| warning: FALSE
#| message: FALSE
#| fig-width: 12
#| fig-height: 8

library(dplyr)
library(ggplot2)
library(sf)
library(rnaturalearth)
library(stringr)
library(stringi)
library(scales)

options(scipen = 999)

il_ozet_out <- final_data_clean %>%
  
  mutate(
    clean_neden = tolower(stri_trans_general(neden, "Latin-ASCII")),
    
    standart_neden = case_when(
      str_detect(clean_neden, "egitim|education") ~ "Education",
      
      str_detect(clean_neden, "konut|yasam|housing|living") ~ "Better Living",
      
      str_detect(clean_neden, "memleket|geri donme|returning") ~ "Returning Home",
      
      str_detect(clean_neden, "bagimli|hane|household|family") ~ "Family Related",
      
      TRUE ~ "Other"
    )
  ) %>%
  
  group_by(il) %>%
  
  mutate(total_out_migration = sum(giden_sayi, na.rm = TRUE)) %>%
  
  filter(giden_sayi == max(giden_sayi, na.rm = TRUE)) %>%
  
  slice(1) %>%
  
  select(
    il,
    top_reason = standart_neden,
    total_out_migration
  ) %>%
  
  ungroup()

clean_city_name <- function(x) {
  x %>%
    as.character() %>%
    stri_trans_general("Latin-ASCII") %>%
    tolower() %>%
    str_trim() %>%
    str_replace_all("ı", "i") %>%
    str_replace_all("^afyon$", "afyonkarahisar") %>%
    str_replace_all("^maras$", "kahramanmaras")
}

turkey_map <- ne_states(
  country = "Turkey",
  returnclass = "sf"
)

map_data_out <- turkey_map %>%
  
  mutate(match_name = clean_city_name(name)) %>%
  
  left_join(
    il_ozet_out %>%
      mutate(match_name = clean_city_name(il)),
    by = "match_name"
  )

map_points_out <- st_point_on_surface(map_data_out)

ggplot() +
  
  geom_sf(
    data = turkey_map,
    fill = "gray95",
    color = "gray70",
    linewidth = 0.25
  ) +
  
  geom_sf(
    data = map_points_out,
    aes(
      size = total_out_migration,
      color = top_reason
    ),
    alpha = 0.85
  ) +
  
  scale_size_area(
    max_size = 14,
    name = "Total Out-Migration",
    labels = comma
  ) +
  
  scale_color_manual(
    values = c(
      "Education" = "#003366",
      "Better Living" = "#2ca25f",
      "Returning Home" = "#f1c40f",
      "Family Related" = "#5dade2",
      "Other" = "gray60"
    ),
    
    na.value = "gray70",
    
    name = "Top Reason for Out-Migration"
  ) +
  
  labs(
    title = "Out-Migration Map of Turkey",
    
    subtitle = "Top reasons for out-migration and total out-migration volumes by province"
  ) +
  
  theme_void() +
  
  theme(
    plot.title = element_text(
      face = "bold",
      size = 20,
      color = "#003366",
      hjust = 0.5
    ),
    
    plot.subtitle = element_text(
      size = 12,
      color = "#555555",
      hjust = 0.5,
      margin = margin(b = 15)
    ),
    
    legend.position = "bottom",
    
    legend.box = "vertical",
    
    legend.title = element_text(
      face = "bold",
      size = 10
    ),
    
    legend.text = element_text(size = 9),
    
    plot.margin = margin(10, 10, 10, 10)
  ) +
  
  guides(
    color = guide_legend(
      ncol = 3,
      title.position = "top"
    ),
    
    size = guide_legend(
      title.position = "top"
    )
  )
```

The map demonstrates that out-migration intensity varies significantly across provinces. Some provinces experience considerably larger population outflows compared to others, which may reflect economic limitations, educational migration, employment opportunities in larger cities, or changing living preferences.

The dominant migration reasons also differ regionally. In some provinces, education emerges as the primary reason for migration, while in others family-related migration or better living conditions play a stronger role.

Overall, the visualization provides insight into the spatial structure of population loss in Turkey and helps illustrate how migration decisions are influenced by both geographical and socio-economic factors.

---

## Pareto Analysis of Migration Flow

Internal migration is not distributed evenly across provinces in Turkey. Some provinces experience significantly higher migration activity due to economic opportunities, educational institutions, industrial development, transportation accessibility, and population density.

In this section, a Pareto analysis is used to examine how migration flow is concentrated across provinces. The Pareto principle, often referred to as the “80/20 rule,” helps identify whether a relatively small number of provinces account for a large proportion of total migration activity.

For this analysis, total migration flow was calculated by combining incoming and outgoing migration values for each province. Provinces were then ranked from highest to lowest migration volume.

In the chart below:
- The navy bars represent total migration flow for each province.
- The yellow cumulative line shows how migration activity accumulates across provinces.
- The secondary axis displays the cumulative percentage distribution.

This visualization helps demonstrate whether migration activity is centralized around a limited number of major provinces or more evenly distributed throughout the country.

```{r}
#| echo: FALSE
#| warning: FALSE
#| message: FALSE
#| fig-width: 13
#| fig-height: 8
#| out-width: 100%

library(dplyr)
library(ggplot2)
library(scales)

options(scipen = 999)

pareto_plot_data <- final_data_clean %>%
  
  group_by(il) %>%
  
  summarise(
    total_flow =
      sum(gelen_sayi, na.rm = TRUE) +
      sum(giden_sayi, na.rm = TRUE),
    
    .groups = "drop"
  ) %>%
  
  arrange(desc(total_flow)) %>%
  
  mutate(
    cumulative_percent =
      cumsum(total_flow) /
      sum(total_flow) * 100,
    
    il = factor(il, levels = il)
  )

ggplot(
  pareto_plot_data,
  aes(x = il, y = total_flow)
) +
  
  geom_col(
    fill = "#003366",
    alpha = 0.95
  ) +
  
  geom_line(
    aes(
      y = cumulative_percent *
        max(total_flow) / 100,
      
      group = 1
    ),
    
    color = "#f1c40f",
    linewidth = 1.2
  ) +
  
  geom_point(
    aes(
      y = cumulative_percent *
        max(total_flow) / 100
    ),
    
    color = "#f1c40f",
    size = 2
  ) +
  
  scale_y_continuous(
    
    name = "Total Migration Flow",
    
    labels = label_number(
      suffix = " K",
      scale = 1e-3
    ),
    
    sec.axis = sec_axis(
      ~ . * 100 / max(pareto_plot_data$total_flow),
      
      name = "Cumulative Percentage (%)",
      
      labels = function(x)
        paste0(round(x), "%")
    ),
    
    expand = expansion(mult = c(0, 0.05))
  ) +
  
  labs(
    title = "Pareto Analysis of Migration Flow",
    
    subtitle = "Provinces ranked by total incoming and outgoing migration volume",
    
    x = "Provinces"
  ) +
  
  theme_minimal(base_family = "sans") +
  
  theme(
    
    plot.background = element_rect(
      fill = "white",
      color = NA
    ),
    
    panel.background = element_rect(
      fill = "white",
      color = NA
    ),
    
    plot.title = element_text(
      face = "bold",
      size = 20,
      color = "#003366"
    ),
    
    plot.subtitle = element_text(
      size = 12,
      color = "#666666",
      margin = margin(b = 15)
    ),
    
    axis.text.x = element_text(
      angle = 90,
      vjust = 0.5,
      hjust = 1,
      size = 7,
      color = "#555555"
    ),
    
    axis.text.y = element_text(
      color = "#555555"
    ),
    
    axis.title = element_text(
      face = "bold",
      color = "#003366"
    ),
    
    panel.grid.minor = element_blank(),
    
    panel.grid.major.x = element_blank(),
    
    panel.grid.major.y = element_line(
      color = "#f0f0f0"
    ),
    
    plot.margin = margin(
      25,
      25,
      25,
      25
    )
  )
```

The Pareto chart clearly shows that migration activity in Turkey is highly concentrated in a relatively small number of provinces. Major metropolitan provinces contribute a disproportionately large share of total migration flow, while many provinces remain at comparatively lower levels.

The cumulative percentage curve rises rapidly during the first part of the distribution, indicating that a limited number of provinces account for a substantial portion of migration activity. This suggests that internal migration in Turkey is strongly centered around economically and socially attractive urban areas.

Another important observation is the existence of strong migration outliers. Provinces such as Istanbul, Ankara, and Izmir stand significantly above the rest in terms of migration flow. These provinces likely function as major hubs for education, employment, transportation, and economic opportunity.

Overall, the Pareto analysis highlights the unequal spatial distribution of migration and demonstrates how migration dynamics are concentrated around a small number of dominant provinces.
