---
title: ""
---

## Data Discovery and Retrieval Process

To meet the project requirement of using a "real-world dataset connected to Türkiye," we selected official statistics from the **Turkish Statistical Institute (TÜİK)**. We focused on the most recent "Internal Migration Statistics" (2024) and specifically utilized the "Inter-provincial migration by reason of migration" tables.

#### Data Source: [TÜİK Data Portal](https://veriportali.tuik.gov.tr/tr/press/54082)

## Dataset Introduction

Our dataset details the migration movements across Türkiye's 81 provinces based on 10 fundamental reasons:

1.  Job Assignment / Change of Job
2.  Starting a Job / Finding a Job
3.  Education
4.  Marital Status / Family-related Reasons
5.  Better Housing and Living Conditions
6.  Migration Dependent on a Household Member
7.  Retirement
8.  Earthquake (Special provision)
9.  Health / Care
10. Other / Unknown

## Data Structure and Exploration

Below, we load our cleaned dataset (`final_data_clean`) and examine its technical structure. This step ensures that our data types (numeric, character, etc.) are correctly assigned for further analysis.

## Libraries and Data Loading

In this section, we load the necessary libraries and our cleaned dataset.

```{r}
#| echo: TRUE
#| warning: FALSE

suppressPackageStartupMessages({
  library(tidyverse)
  library(readr)
  library(knitr)
})

options(encoding = "UTF-8")

final_data_clean <- read_csv(
  "final_data_clean.csv",
  locale = locale(encoding = "UTF-8")
)
```

## Structure of Dataset

```{r}
#| echo: TRUE

str(final_data_clean)
```

## Data Frame

```{r}
#| echo: FALSE

cat("Data Frame:", "\n")
cat("Number of Observations:", nrow(final_data_clean), "\n")
cat("Number of Variables:", ncol(final_data_clean), "\n\n")

cat("Variable Names:", "\n")
print(colnames(final_data_clean))
```

## Migration by Province

### Top 20 Incoming Migration by Province

```{r}
#| echo: TRUE
#| fig-width: 10
#| fig-height: 8

library(dplyr)
library(ggplot2)
library(scales)

final_data_clean %>%
  group_by(il) %>%
  summarise(total_incoming = sum(gelen_sayi, na.rm = TRUE)) %>%
  arrange(desc(total_incoming)) %>%
  slice_head(n = 20) %>%
  ggplot(aes(x = reorder(il, total_incoming), y = total_incoming)) +
  geom_col(fill = "#001f5b", width = 0.75) +
  coord_flip() +
  scale_y_continuous(labels = comma) +
  labs(
    title = "Top 20 Provinces by Incoming Migration",
    x = NULL,
    y = "Incoming Migration"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(size = 18, face = "bold"),
    axis.text.y = element_text(size = 11),
    axis.text.x = element_text(size = 10),
    panel.grid.major.y = element_blank(),
    panel.grid.minor = element_blank()
  )
```

### Top 20 Outgoing Migration by Province

```{r}
#| echo: TRUE
#| fig-width: 10
#| fig-height: 8

final_data_clean %>%
  group_by(il) %>%
  summarise(total_outgoing = sum(giden_sayi, na.rm = TRUE)) %>%
  arrange(desc(total_outgoing)) %>%
  slice_head(n = 20) %>%
  ggplot(aes(x = reorder(il, total_outgoing), y = total_outgoing)) +
  geom_col(fill = "#001f5b", width = 0.75) +
  coord_flip() +
  scale_y_continuous(labels = comma) +
  labs(
    title = "Top 20 Provinces by Outgoing Migration",
    x = NULL,
    y = "Outgoing Migration"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(size = 18, face = "bold"),
    axis.text.y = element_text(size = 11),
    axis.text.x = element_text(size = 10),
    panel.grid.major.y = element_blank(),
    panel.grid.minor = element_blank()
  )
```

## Most Common Reasons for Migration

```{r}
#| echo: TRUE
#| fig-width: 11
#| fig-height: 8

library(dplyr)
library(ggplot2)
library(scales)

final_data_clean %>%
  group_by(neden) %>%
  summarise(total_migration = sum(gelen_sayi + giden_sayi, na.rm = TRUE)) %>%
  arrange(desc(total_migration)) %>%
  slice_head(n = 10) %>%
  ggplot(aes(x = reorder(neden, total_migration),
             y = total_migration)) +
  geom_col(fill = "#001f5b", width = 0.75) +
  coord_flip() +
  scale_y_continuous(labels = comma) +
  labs(
    title = "Top 10 Reasons for Migration",
    x = NULL,
    y = "Total Migration"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(size = 18, face = "bold"),
    axis.text.y = element_text(size = 10),
    axis.text.x = element_text(size = 10),
    panel.grid.major.y = element_blank(),
    panel.grid.minor = element_blank()
  )
```
